---
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include = FALSE}
library(ggplot2)
library(data.table)
library(sqldf)

#Loading tables
setwd('C:/Users/SHARON/Desktop/R/Data Science Project 1/Proj1_4')
sqlite    <- dbDriver("SQLite")
con <- dbConnect(sqlite, dbname='R1_4.SQLite')
dbBegin(con)

LAF_MoTY <- data.table(dbReadTable(con, 'LAF_MoTy'))
LRF_MoTY <- data.table(dbReadTable(con, 'LRF_MoTy'))
LDF_MoTY <- data.table(dbReadTable(con, 'LDF_MoTy'))

LAF_Year <- data.table(dbReadTable(con, 'LAF_Year'))
LRF_Year <- data.table(dbReadTable(con, 'LRF_Year'))
LDF_Year <- data.table(dbReadTable(con, 'LDF_Year'))

LAFmap <- data.table(dbReadTable(con, 'LAFmap'))
LRFmap <- data.table(dbReadTable(con, 'LRFmap'))
LDFmap <- data.table(dbReadTable(con, 'LDFmap'))

LALS_MoTY <- data.table(dbReadTable(con, 'LALS_MoTy'))
LRLS_MoTY <- data.table(dbReadTable(con, 'LRLS_MoTy'))
LDLS_MoTY <- data.table(dbReadTable(con, 'LDLS_MoTy'))

LALSmap <- data.table(dbReadTable(con, 'LALSmap'))
LRLSmap <- data.table(dbReadTable(con, 'LRLSmap'))
LDLSmap <- data.table(dbReadTable(con, 'LDLSmap'))

LADTI_MoTY <- data.table(dbReadTable(con, 'LADTI_MoTY'))
LRDTI_MoTY <- data.table(dbReadTable(con, 'LRDTI_MoTY'))
LDDTI_MoTY <- data.table(dbReadTable(con, 'LDDTI_MoTY'))

LADTImap <- data.table(dbReadTable(con, 'LADTImap'))
LRDTImap <- data.table(dbReadTable(con, 'LRDTImap'))
LDDTImap <- data.table(dbReadTable(con, 'LDDTImap'))

LATmap <- data.table(dbReadTable(con, 'LATmap'))
LRTmap <- data.table(dbReadTable(con, 'LRTmap'))

```

#Loan Summary Statistics


###Introduction
The goal of this project is to create a set of charts that examine who, when, and where the ideal conditions are for lending club to loan money to an individual. With a large
Company like lending club there are a multitude of various factors at play from what
state does the individual reside in to how many loans they have already given out in 
a given period. Since we are observing from an outside perspective of the company we have
no way of being sure if there are quotas set for employees or other nuances to this
companies particular business model, so it will be interesting to examine these factors
in depth.

in this report we will be using 3 types of charts, Acception vs rejection, a chart with
a pair of lines that shows the data from loans that were accepted against loans that
were rejected; percentage difference, a chart that is ment to quantify the differences
in terms of percentages between the two groups and to make it easier to see the
differences between the two; and heat maps, to show differences over geographical regions
in our case the lower 48 states.


###Loan Frequency vs. Months of the Year:
Frequency plots based on twelve months of the year, and number of loans taken out during each month. months with less than 31 days we found the mean frequency per day, scaled it to 31 days and then normalized them based on the total frequency under Loans accepted and Loans Rejected

```{r}
ggplot()+ 
geom_line(data=LAF_MoTY, aes(x=V1, y=Norm, group=1, colour='Accepted Loan'), color='green')+
geom_line(data=LRF_MoTY, aes(x=V1, y=Norm, group=1, colour='Rejected Loan'), color='red')+
  xlab('Months of the Year') + ylab('Normalized Frequency') +
  ggtitle('Loan request Frequency vs Months of the Year')
```

In the chart above, the green line represents the Normalized number of loans accepted per
month while the red line represents the normalized number of loans rejected per month.

For a consumer interested in taking a loan out from lending club we can see that september
has the highest frequency of loans accepted out of any month, while august on the other hand has the lowest number of loans accepted per month.



```{r}
ggplot()+ 
geom_line(data=LDF_MoTY, aes(x=V1, y=Norm, group=1, color='Accepted Loan'),color='blue')+
  xlab('Months of the Year') + ylab('Normalized Frequency Ratio') +
  ggtitle('Loan Ratio Frequency vs Months of the Year')
```

The above chart is created to confirm our hypothesis from the two lines in the first chart
on Loan frequency. In september there is roughly a 20% increase in loan acceptances
over loan rejections over all. and in total we can see there are 7 / 12 months that
have a higher rate of loans being accepted than loans being rejected.


###Percent Change in Loan Frequency vs. Business Year:
Loan Frequency by Year;
These group of charts plots the percentage increase in the # of loans from the previous year.

```{r}
ggplot()+ 
geom_line(data=LAF_Year, aes(x=V1, y=N, group=1, color='Accepted Loan'), color='green')+
geom_line(data=LRF_Year, aes(x=V1, y=N, group=1, color='Rejected Loan'), color='red')+
  xlab('Years in Business') + ylab('Percent Change from Previous Year')+
  ggtitle('Loan Frequency Changes vs Years in Business')
```

In the above chart the green line represents the percent increase in number of loans
accepted over the previous year, and the red line represents the number of loans rejected 
over the previous year.

In years where the % increase in number of loans accepted is higher then the number of
loans rejected we could conclude that the lending club business is doing very well since
it must be growing to take out those loans, and therefore would be more likely to accept
a loan than other years.

An interesting note with the above is the extremely high number of loans rejected between
the years 2009 - 2011, which are considered some of the worst years in the great
recession.
```{r}
ggplot()+ 
geom_line(data=LDF_Year, aes(x=V1, y=N, group=1, color='Accepted Loan'), color='blue')+
  xlab('Years in Business') + ylab('Percent Change from Previous Year')+
  ggtitle('Loan Frequency Ratio vs Years in Business')
  
```

The above chart represents the percent growth in accepted loans divided the percent growth
in rejected loans. We can see that at its worst lending club increased its number of
accepted loans at a rate of 40% compared to its rejections. 2010 was not the year to be
a consumer taking a loan from lending club. looking towards the present we can see that 
the fluctuations have gone down a lot and averages roughly around 80% change in loans 
accepted vs loans rejected. However, arguements can be made that this is not entirely due
to lending club being unsuccessful as before, as it could also be attributed to a healthy
economy. During the recession many people would have needed small loans to pay bills or
make purchases. But, since around 2012, unemployment rate has been steadly declining 
meaning more people can make purchases or have savings developed leading to less loans
being applied for.

###Relative Frequency of Loans vs. Lower Fourty-Eight States:

Normalized over two values to account for # of loans requested and State population
Size.
```{r}
LAFmap <- LAFmap[order(LAFmap$order),]
ggplot(data = LAFmap, aes(x=long, y=lat, group = group)) +
  geom_polygon(aes(fill = N )) + geom_path() +
  scale_fill_gradientn(colours=rev(heat.colors(10)),na.value="grey10") + coord_map()+
  xlab('Latitude') + ylab('Longitude')+
  ggtitle('Loans Accepted Frequency vs State')
  
```

```{r}
LRFmap <- LRFmap[order(LRFmap$order),]
ggplot(data = LRFmap, aes(x=long, y=lat, group = group)) +
      geom_polygon(aes(fill = N)) + geom_path() +
      scale_fill_gradientn(colours=rev(heat.colors(10)),na.value="grey10") + coord_map()+
  xlab('Latitude') + ylab('Longitude')+
  ggtitle('Loans Rejected Frequency vs State')

```

This final plot out of the three below is a ratio of number of loans accepted vs number
of loans rejected. 

We can see that lending club tends to reject more loans from south eastern united states
while accepting more loans towards the north east and west coast areas.

```{r}
LDFmap <- LDFmap[order(LDFmap$order),]
ggplot(data = LDFmap, aes(x=long, y=lat, group = group)) +
      geom_polygon(aes(fill = N)) + geom_path() +
      scale_fill_gradientn(colours=rev(heat.colors(10)),na.value="grey10") + coord_map()+
  xlab('Latitude') + ylab('Longitude')+
  ggtitle('Loans Frequency Ratio vs State')

```

###Average Loan Size vs. Month of the Year:
This group of plots is created to show the average Loan sizes accepted and rejected
Based on the month of the year.

Across the board we can see larger loans tend to be accepted while smaller loans
tend to be rejected. However, if a consumer is interested in taking  out a smaller loan
the best time to do so is towards the end of the year, as lending clubs average
rejected loan size decreasese continually as it approaches towards december

```{r}
ggplot()+ 
geom_line(data=LALS_MoTY, aes(x=AppMo, y=V1, group=1, color='Accepted Loan'), color='green')+
geom_line(data=LRLS_MoTY, aes(x=AppMo, y=V1, group=1, color='Rejected Loan'), color='red')+
  xlab('Months of the Year') + ylab('Average Loan Size')+
  ggtitle('Average Loan Size vs Month of the Year')

```

This plot below is the ratio between loan size accepted and loan size rejected per month.
This clearly shows the trend we could see in our first charts, that as we approach the 
end of the year the average rejected loan size decreases faster than the changes in 
average accepted loan size.

```{r}
ggplot()+ 
geom_line(data=LDLS_MoTY, aes(x=AppMo, y=V1, group=1, color='Accepted Loan'), color='blue')+
  xlab('Months of the Year') + ylab('Average Loan Size')+
  ggtitle('Average Loan Size Ratio vs Month of the Year')
```


###Loan Size vs. Lower Fourty-Eight States:
Below is a set of heat map plots over the lower fourty-eight united states showing
the average loan sizes of each state accepted and rejected. The interesting part of these
plots is Iowa, Iowa is the only state with an average accepted loan size lower than
its average rejected loan size, which is why our first chart looks so dark.

```{r}
LALSmap <- LALSmap[order(LALSmap$order),]
ggplot(data = LALSmap, aes(x=long, y=lat, group = group)) +
      geom_polygon(aes(fill = V1)) + geom_path() +
      scale_fill_gradientn(colours=rev(heat.colors(30)),na.value="grey40") + coord_map()+
  xlab('Latitude') + ylab('Longitude')+
  ggtitle('Average Loan Accepted Size vs State')

```

```{r}
LRLSmap <- LRLSmap[order(LRLSmap$order),]
ggplot(data = LRLSmap, aes(x=long, y=lat, group = group)) +
      geom_polygon(aes(fill = V1)) + geom_path() +
      scale_fill_gradientn(colours=rev(heat.colors(10)),na.value="grey10") + coord_map()+
  xlab('Latitude') + ylab('Longitude')+
  ggtitle('Average Loan Rejected Size vs State')
```

The plot below is the best one to examine this peice of data. If you're taking out a 
small loan the best place to go is Iowa. However, if you are taking out a large loan
we can see the best place to do so is the south east and Texas, with florida being 
the only exception.

```{r}
LDLSmap <- LDLSmap[order(LDLSmap$order),]
ggplot(data = LDLSmap, aes(x=long, y=lat, group = group)) +
      geom_polygon(aes(fill = V1)) + geom_path() +
      scale_fill_gradientn(colours=rev(heat.colors(48)),na.value="grey10") + coord_map()+
  xlab('Latitude') + ylab('Longitude')+
  ggtitle('Average Loan Size Ratio vs State')
```

###Average Consumer Debt to Income Ratio vs Month of the Year:
These group of plots show the average accepted and rejected loan's debt to income ratio
at the time the consumer applied for it. we can see that unlike loan size higher debt
to income ratios are rejected while lower ones are accepted.
```{r}
ggplot()+ 
geom_line(data=LADTI_MoTY, aes(x = AppMo, y=V1, group=1, color='Accepted Loan'), color='green') +
geom_line(data=LRDTI_MoTY, aes(x=AppMo, y=V1, group=1, color='Rejected Loan'), color='red') +
  xlab('Months of the Year') + ylab('Average Consumer Debt to Income Ratio')+
  ggtitle('Average Loan Debt to Income Ratio vs Months of the Year')
```

An interesting trend we can spot here is again as we approach december the ratio between
accepted loan debt to income ratio and rejected loan debt to income ratio goes up 
dramatically. This shows a clear trend that as we approach the later months of the year
lending club relaxes their requirments for taking out a loan.

```{r}
ggplot()+ 
geom_line(data=LDDTI_MoTY, aes(x=AppMo, y=V1, group=1, color='Accepted Loan'), color='blue')+
   xlab('Months of the Year') + ylab('Average Consumer Debt to Income Ratio')+
  ggtitle('Average Debt to Income Ratio vs Months of the Year ')
```

###Average Consumer Debt to Income Ratio vs Lower Fourty-Eight States:
Accepted and Rejected loans debt to income ratio vs lower 48 states. For consumers
states with the easiest loans to get would be ones that have higher debt to income ratios
on accepted loans and higher debt to income ratios on rejected loans. Below we can see
the areas where this occurs is mostly in the mid-western united states.

```{r}
LADTImap <- LADTImap[order(LADTImap$order),]
ggplot(data = LADTImap, aes(x=long, y=lat, group = group)) +
      geom_polygon(aes(fill = V1)) + geom_path() +
      scale_fill_gradientn(colours=rev(heat.colors(10)),na.value="grey10") + coord_map()+
    xlab('Latitude') + ylab('Longitude')+
  ggtitle('Average Accepted Loan Debt to Income Ratio vs State')
```

```{r}
LRDTImap <- LRDTImap[order(LRDTImap$order),]
ggplot(data = LRDTImap, aes(x=long, y=lat, group = group)) +
      geom_polygon(aes(fill = V1)) + geom_path() +
      scale_fill_gradientn(colours=rev(heat.colors(10)),na.value="grey10") + coord_map()+
  xlab('Latitude') + ylab('Longitude')+
  ggtitle('Average Rejected Loan Debt to Income Ratio vs State')
```


For our ratio plot, a larger ratio between loans accepted and loans rejected is ideal for
consumers since it implies that even with an average loan accepted DTI, if the state 
average for rejection is higher that consumers DTI could still be adequate. 
```{r}
LDDTImap <- LDDTImap[order(LDDTImap$order),]
ggplot(data = LDDTImap, aes(x=long, y=lat, group = group)) +
      geom_polygon(aes(fill = V1)) + geom_path() +
      scale_fill_gradientn(colours=rev(heat.colors(10)),na.value="grey10") + coord_map()+
  xlab('Latitude') + ylab('Longitude')+
  ggtitle('Average Debt to Income Ratio vs Lower Fourty-Eight States')
```

###Loan Title Frequency vs Lower Fourty-Eight States:
Frequency of loan purpose vs lower 48 states. In the last two plots below we've taken 
the frequency counts of all the accepted loans titles and all the rejected loans titles
and checked which state had the highest of each. 

For accepted loans we can see that the most frequently occuring reason is medical. Its
good to know that Lending Club is helping people out in their times of need and as a 
consumer it must feel reassuring to hear.



```{r}
LATmap <- LATmap[order(LATmap$order),]
ggplot(data = LATmap, aes(x=long, y=lat, group = group)) +
      geom_polygon(aes(fill = Title)) + geom_path() + coord_map()+
    xlab('Latitude') + ylab('Longitude')+
  ggtitle('Highest Reason for Loan Acceptance per State')
```

Looking at rejected loans, they're all over the place but it seems that student loans
are the most commonly rejected reason. However, student loans outside of the goverment 
are very risky for investors since unlike with material objects or medical bills
student loans don't guarantee that person will have a job or be able to pay these bills
back. Further with government student loans there is even less risk since wages 
can be garnished if you declare bankruptcy, but the same is not true for the private
sector.

```{r}
LRTmap <- LRTmap[order(LRTmap$order),]
ggplot(data = LRTmap, aes(x=long, y=lat, group = group)) +
      geom_polygon(aes(fill = Title)) + geom_path() + coord_map()+
    xlab('Latitude') + ylab('Longitude')+
  ggtitle('Highest Reason for Loan Rejection vs State')

```


###Conclusions:
After examining through all of our data from above, I believe we can come to several 
interesting conclusions. By examining the average consumers debt to income ratio and loan sizes vs months of the year, we can see a trend as the end of the year approaches where
the number of loans rejected drops dramatically. From this we can assume that as the
years end approaches Lending Club sales people are trying to meet end of the year quota
so less satisfactory loans will be accepted to make ends meet. Another assumption we can 
make is during the last few months of the year the holiday season is here. Many people
are Christmas shopping and therefore need smaller loans to make purchases for friends
and family. While some people would still get rejected the number of people who are
looking for a small loan go from people who need a payday loan to all sorts of folks.

As for locational trends, We can notice two additional trends. As we look towards the 
south eastern part of the united states the ratio between accepted loan size and rejected
loan size is the largest in the country, and concentrated to mostly this area. The other 
trend we see comes from consumer debt to income ratio, in the mid west we can see has the
largest differences between accepted loans debt to income ratio and rejected loans debt 
to income ratio. Both of these can be understood by examining the local populations
closer. For the south-eastern united states we know on average have lower wages and
would tend to request smaller loan sizes, further if a higher percentage of the 
population tends to live closer to the poverty line, they would have other outstanding
reasons as to why even though they have some of the smallest loan rejected sizes they
have some of the highest percentage of loan rejections vs state population. Looking at 
the mid-west with and its trend in consumer debt to income ratio, its the oppisite.
while there is still higher instances of people living closer to the poverty line out 
in this region there are still very few people living in these areas vs the larger 
coastal cities. For Lending Club to get its business out there and have consumers hear 
about them, they would need to lower the requirements for a loan to get accepted so 
they can spread their business through word of mouth.